---
title: "India_Population"
author: "Jigyasa Bhagat"
date: "March 9, 2016"
output: 
  html_document: 
    keep_md: yes
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
# devtools::install_github('walkerke/idbr')
# install.ImageMagick(URL = "http://www.imagemagick.org/script/binary-releases.php")
# install.packages("animation",dep=T)
library(animation)
library(idbr)
library(ggplot2)
library(dplyr)
library(ggthemes)
# require("installr")
idb_api_key("xxxxxxxxxxxxxxxxxxxxxxxxxx")
```
This code and image is inspired by http://blog.revolutionanalytics.com/2016/02/japans-ageing-population-animated-with-r.html

```{r, include=T,message=F}
male=idb1('IN',2000:2050,sex="male") %>%
    mutate(POP=POP * -1,
           SEX='Male')
female=idb1('IN',2000:2050,sex="female") %>% mutate(SEX="Female")
india=rbind(male,female) %>% mutate(abs_pop=abs(POP))
```

# Reproducing Plots. 
In addition to reproducing a similar plot for India for the year 2000 to 2050, I added median age for Male and Female. (For each year, which age had maximum population). Length of the meadian age segment is equal to the population at the age for that year.
```{r pressure, include=T,message=F,warning=F}
saveGIF({
    for (i in 2000:2050){
        title<-as.character(i)
        year_data<-filter(india,time==i)
        median_age_male_tmp=filter(year_data,SEX=="Male")
        median_age_male=median_age_male_tmp$AGE[which.max(median_age_male_tmp$abs_pop)]
        median_age_female_tmp=filter(year_data,SEX=="Female")
        median_age_female=median_age_female_tmp$AGE[which.max(median_age_female_tmp$abs_pop)] 
    g1<-ggplot(year_data,aes(x=AGE,y=POP,fill=SEX,width=1))+
        coord_fixed()+
        coord_flip()+
        annotate('text',x=98,y=-1e+07,
            label='Data: US Census Bureau IDB; idbr R package; Jigyasa Bhagat',size=3)+
         geom_bar(data=subset(year_data,SEX=="Female"),stat="identity")+
        geom_bar(data=subset(year_data,SEX=="Male"),stat="identity")+
        geom_segment(aes(x=median_age_male,xend=median_age_male,
                         y=year_data$POP[year_data$AGE==median_age_male & year_data$SEX=="Male"],yend=0))+
        geom_segment(aes(x=median_age_female,xend=median_age_female,
                    y=year_data$POP[year_data$AGE==median_age_female & year_data$SEX=="Female"],yend=0))+
        scale_y_continuous(breaks=seq(-1e+07, 1e+07, 5e+06),
            labels=paste0(as.character(c(seq(10, 0, -5), c(5, 10))),"m"),
            limits=c(min(india$POP),max(india$POP))) +
        theme_economist(base_size = 14) +
        scale_fill_manual(values = c('#D07371','#5BC5C8')) +
        ggtitle(paste0('Population structure of India, ', title)) +
        ylab('Population') +
        xlab('Age') +
        theme(legend.position = "bottom", legend.title = element_blank())+
        guides(fill = guide_legend(reverse = TRUE))
      
    print(g1)     
    }
},movie.name='india_pyramid.gif',interval=0.2,ani.width=700,ani.height=600)
```




# Sex ratio visualization for years 2000 to 2050
This visualization represents the sex ratio (Male / Female) for all ages (0 to 100) for the years 2000 to 2050
```{r,include=T,message=F,warning=F}
india_ratio=mutate(male,ratio=abs(male$POP/female$POP))
cl=rainbow(50)
saveGIF({
    for(i in 2000:2050){
        x<-india_ratio$ratio
        if (i<2050)
                x[india_ratio$time>i]<-NA
        plot(india_ratio$AGE,x,ylim=c(0,1.20),main=paste("India:Sex Ratio(",i,")",sep=""),type="n",xlab = "AGE", ylab = "Number of Males/Number of Females")
        for (j in 2000:i){
            lines(india_ratio$AGE[india_ratio$time==j],x[india_ratio$time==j],col=ifelse(j==2000,"red",cl[as.numeric(substr(j, 3, 4))]),lwd=1.5)
        }
            
        if (i>2000)
                x[india_ratio$time<i]<-NA
        lines(india_ratio$AGE,x,col=1,lwd=2)
}
},movie.name='india_male_female_ratio.gif',interval=0.2,ani.width=700,ani.height=600)
```

